The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-26 formalwriteup.tex
Add comment on generic c

\begin{corollary}\label{corr:dFdz2}
Let $s_1,\dots,s_{r}$ be distinct elements of $k$ and $m_i$ positive integers such that $\sum_i m_is_i=0$.  
For some $V(z)$ satisfying $[z^l]V(z)=0$ for $0 \le l \le p-1$, we have
\[
F'(z)=V(z) - \sum_j \frac{cm_js_j}{1-s_jz}F(z).
\]
\end{corollary}
\begin{proof}
Apply Lemma~\ref{lem:dFdz} by letting the $x_i$ be $m_i$ copies of the $s_i$. 
\end{proof}

\begin{lemma} \label{lem:resbyparts}
Let $s_1,\dots,s_{r}$ be distinct non-zero elements of $k$ and $m_i$ positive integers so that that $\sum_i m_is_i=0$.  For any integer $l \le p-r$, $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\ldots$, $[z^{l+r}]a(z)$.
\end{lemma}
\begin{proof} 
We will use an analogue of ``residues by parts'' in a formal power series setting.   We first show that 
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)s_j}{1-s_jz}a(z)
\]
for some formal power series $b(z)$. By Corollary~\ref{corr:dFdz2} we have for some $V(z)$ where $[z^l]V(z)=0$ for $0 \le l \le p-1$ that $F'(z)=V(z)+\sum_j \frac{-cm_js_j}{1-s_jz}F(z)$. We now compute
\begin{align*}
a'(z)=\frac{d}{dz}\left(\frac{1}{\prod_{i=1}^{r}(1-s_iz)}F(z)\right)&=\frac{1}{\prod_{i=1}^{r}(1-s_iz)}F'(z)+\sum_{j=1}^{r}\frac{s_j}{(1-s_jz)\prod_{i=1}^{r}(1-s_iz)}F(z)\\
&=\frac{1}{\prod_{i=1}^{r}(1-s_iz)}\left(V(z)+\sum_j \frac{-cm_js_j}{1-s_jz}F(z)\right)+\sum_{j=1}^{r}\frac{s_j}{1-s_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=1}^{r}(1-s_iz)}-\sum_j \frac{(m_jc-1)s_j}{1-s_jz}a(z).
\end{align*}
The fact that $[z^l]V(z)=0$ for $0 \le l \le p-1$ then gives the claim.

@@

We first claim that at least one $s_i$ is $0$. Suppose for the sake of contradiction that $\prod_i s_i \ne 0$.  We recall
\[
a(z)=\frac{1}{\prod_{i=1}^{r}(1-s_iz)}F(z)
\]
as in Lemma~\ref{lem:resbyparts}.  Recall that $b_i(z)=\prod_{j \ne i} (1-s_jz)$ for $i=1,\dots,r$.   Now, for any $i$, we have $a(z)b_i(z)=\frac{1}{1-s_iz}F(z)$ so in particular 
\[
[z^p]a(z)b_i(z)=[z^p]\frac{1}{1-s_iz}F(z)=0.
\]
Therefore for any $\lambda_i \in k$, we have
\begin{equation}\label{eq:lincomb}
\sum_{i=1}^{r} \lambda_i[z^p]a(z)b_i(z)=[z^p] \,a(z)\sum_{i=1}^{r}\lambda_ib_i(z)=0.
\end{equation}
By Lemma~\ref{lem:vand}, $b_i(z)$ are linearly independent polynomials of degree at most $r-1$. Therefore for $0 \le m \le r-1$ we can choose $\{\lambda_i^m\}$ so that $\sum_i\lambda_i^mb_i(z)=z^m$. Choosing these $\lambda_i^m$ in (\ref{eq:lincomb}), we see that
\[
[z^{p-r+1}]a(z)=\dots=[z^p]a(z)=0.
\]
By repeated application of Lemma~\ref{lem:resbyparts} starting with $l=p-r$, we find that 
\[
[z^l]a(z)=0 \text{ for $l=p-r,p-r-1,\dots,0$}.
\]
This implies that $[z^0]a(z)=0$, a contradiction. We conclude that $s_i=0$ for some $i$.

>>>>

\begin{corollary}\label{corr:dFdz2}
Let $s_1,\dots,s_{r}$ be distinct elements of $k$ and $m_i$ positive integers such that $\sum_i m_is_i=0$.  
For some $V(z)$ satisfying $[z^l]V(z)=0$ for $0 \le l \le p-1$, we have
\[
F'(z)=V(z) - \sum_j \frac{cm_js_j}{1-s_jz}F(z).
\]
\end{corollary}
\begin{proof}
Take the $x_i$ to be $m_i$ copies of the $s_i$ in Lemma~\ref{lem:dFdz}.
\end{proof}

\begin{lemma} \label{lem:resbyparts}
Let $s_1,\dots,s_{r}$ be distinct non-zero elements of $k$ and $m_i$ positive integers so that that $\sum_i m_is_i=0$.  For generic $c$ and any integer $l \le p-r$, $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\ldots$, $[z^{l+r}]a(z)$.
\end{lemma}
\begin{proof} 
We will use an analogue of ``residues by parts'' in a formal power series setting.   We first show that 
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)s_j}{1-s_jz}a(z)
\]
for some formal power series $b(z)$. By Corollary~\ref{corr:dFdz2} we have for some $V(z)$ where $[z^l]V(z)=0$ for $0 \le l \le p-1$ that $F'(z)=V(z)+\sum_j \frac{-cm_js_j}{1-s_jz}F(z)$. We now compute
\begin{align*}
a'(z)=\frac{d}{dz}\left(\frac{1}{\prod_{i=1}^{r}(1-s_iz)}F(z)\right)&=\frac{1}{\prod_{i=1}^{r}(1-s_iz)}F'(z)+\sum_{j=1}^{r}\frac{s_j}{(1-s_jz)\prod_{i=1}^{r}(1-s_iz)}F(z)\\
&=\frac{1}{\prod_{i=1}^{r}(1-s_iz)}\left(V(z)+\sum_j \frac{-cm_js_j}{1-s_jz}F(z)\right)+\sum_{j=1}^{r}\frac{s_j}{1-s_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=1}^{r}(1-s_iz)}-\sum_j \frac{(m_jc-1)s_j}{1-s_jz}a(z).
\end{align*}
The fact that $[z^l]V(z)=0$ for $0 \le l \le p-1$ then gives the claim.

@@

We first claim that at least one $s_i$ is $0$. Suppose for the sake of contradiction that $\prod_i s_i \ne 0$.  We recall
\[
a(z)=\frac{1}{\prod_{i=1}^{r}(1-s_iz)}F(z)
\]
as in Lemma~\ref{lem:resbyparts}.  Recall that $b_i(z)=\prod_{j \ne i} (1-s_jz)$ for $i=1,\dots,r$.   Now, for any $i$, we have $a(z)b_i(z)=\frac{1}{1-s_iz}F(z)$ so in particular 
\[
[z^p]a(z)b_i(z)=[z^p]\frac{1}{1-s_iz}F(z)=0.
\]
Therefore for any $\lambda_i \in k$, we have
\begin{equation}\label{eq:lincomb}
\sum_{i=1}^{r} \lambda_i[z^p]a(z)b_i(z)=[z^p] \,a(z)\sum_{i=1}^{r}\lambda_ib_i(z)=0.
\end{equation}
By Lemma~\ref{lem:vand}, $b_i(z)$ are linearly independent polynomials of degree at most $r-1$. Therefore for $0 \le m \le r-1$ we can choose $\{\lambda_i^m\}$ so that $\sum_i\lambda_i^mb_i(z)=z^m$. Choosing these $\lambda_i^m$ in (\ref{eq:lincomb}), we see that
\[
[z^{p-r+1}]a(z)=\dots=[z^p]a(z)=0.
\]
By repeated application of Lemma~\ref{lem:resbyparts} starting with $l=p-r$ (but finitely many times and with finitely many choices of $l$, $r$, and $m_i$), we find for generic $c$ that
\[
[z^l]a(z)=0 \text{ for $l=p-r,p-r-1,\dots,0$}.
\]
This implies that $[z^0]a(z)=0$, a contradiction. We conclude that $s_i=0$ for some $i$.
